---
title: "Day13-Communities to metacommunities"
author: "Katie Barry"
date: "`r Sys.Date()`"
output: html_document
---
**LEARNING GOALS** We have two major learning goals for today: 
1. Scale up our community into a metacommunity
2. Understand the consequences of connection within a metacommunity
3. Brainstorm ways in which we can make out meta-community more realistic 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(primer)
source("./functions/lv.metacommunity.function.R")
```

### Part 1: Reviewing models of communities 

Let's review our model of interspecific competition for a moment. 

Each species has a starting value (y0) each species  the model runs for a certain number of timesteps (t). 
Each species also has an r which is their growth rate and then two alphas, one for their own self-limitation and one for how much they are limited by the other species.

```{r modeling competition between two species in R}
#starting both populations at 1 individuals
y0<-c(1,1)

#the time points that the model will run for
t<-0:100

#setting the parameters
p<-c(r1=1, r2=1, a11=0.2, a12=0.1, a21=0.1, a22=0.2)


out <- ode(y=y0, times=t, func=lvcomp2, parms=p)

plot(out)
```

In this scenario - both species can survive!! Woo hoo! 
*Review questions*
1. What happens when you change the alphas? 
2. What happens when you change r? 


### Part 2: Adding in a metacommunity 

Let's review the concept of meta...

A metapopulation is a group of connected populations. Populations have only one species. 

If we extrapolate this idea to communities, a metacommunity is a group of connected communities! 

That means that in addition to our birth rate and our interaction coefficients (alpha11, alpha12, etc) each species also is embedded in it's own meta-population and therefore can have immigration and emigration. 

For this we use a function that I wrote, it has the same parameters as our interspecific interaction model (above) with the addition of: 
ci1 - the immigration of species 1
ci2 - the immigration of species 2
e1- the emigration of species 1 
e2 - the emigration of species 2
```{r}
#starting both populations at 1 individuals
y0<-c(1,1)

#the time points that the model will run for
t<-0:100

#setting the parameters
p<-c(r1=1, r2=1, a11=0.2, a12=0.1, a21=0.1, a22=0.2, ci1=1, ci2=1, e1=1, e2=1)

#note that I have changed the func to "lv.metacommunity" this is the function that I have written to show you a meta-community model
out <- ode(y=y0, times=t, func=lv.metacommunity, parms=p)

plot(out)
```

What do you notice about these plots? How do they compare from our earlier versions of interspecific interactions? 

Play around with the values for ci1 and ci2, e1 and e2 and answer the following questions. 

*Questions*
1. What happens when ci1 is low but ci2 is high? (hint - try changing ci2 to 10)
2. What happens when e1 is low but e2 is high? (hint-try changing e2 to 10)
3. How do these changes compare to changes in r or one of our alphas? 

### Part 3: Thinking about realism

In the scenarios that I've described - all immigration and emigration are coming and going from **the ether**. They don't have a source. We don't see what's going on in the space around the species. Brainstorm 5 ways that we can make this situation closer to the reality that you see outside your window. 



